---
title: "W271 Group Lab 1"
subtitle: "Investigating the 1986 Space Shuttle Challenger Accident"
author: "Elizabeth, Emily, Michael, Olivia and Srila"
output: bookdown::pdf_document2
fontsize: 11pt
geometry: margin=1in
---

\newpage

```{=tex}
\begin{abstract} 
This report will, indeed, be abstract. No, instead, describe your goals your approach, and what you learn.
\end{abstract}
```

```{r load packages, message=FALSE, warning=FALSE, echo=FALSE}
list.of.packages <- c("tidyverse", 
                      "sandwich", 
                      "lmtest", 
                      "car", 
                      "MASS", 
                      "nnet", 
                      "cowplot")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(tidyverse)
library(sandwich)
library(lmtest)
library(car)
library(MASS)
library(nnet)
library(cowplot)
library(gridExtra)
library(corrplot)
```

# Introduction

## Research question

# Data (20 points)

**Complete the following task. In your final submission, please remove this 
question prompt so that your report reads as a report. The Data Section of this 
report is worth 20 points.**

-   Conduct a thorough EDA of the data set.

    -   This should include both graphical and tabular analysis as taught in 
    this course.
    -   Since the report has a page-limit, you will have to be selective when 
    choosing visuals to illustrate your key points, associated with a concise 
    explanation of the visuals.

-   This EDA should begin with an inspection of the given dataset; examination 
of anomalies, missing values, potential of top and/or bottom code etc.

```{r, EDA}
df <- read_csv('/Users/srilamaiti/Downloads/data/raw/challenger.csv')
summary(df)
head(df)
tail(df)
str(df)
dim(df)
which(is.na(df$Flight))
which(is.na(df$Temp))
which(is.na(df$Pressure))
which(is.na(df$O.ring))
which(is.na(df$Number))
unique(df$Flight)
unique(df$Temp)
unique(df$O.ring)
unique(df$Number)

df %>%
  count(O.ring) %>%
  mutate(prop = round(prop.table(n),2)) %>%
  kable(col.names = c('O.ring_count', 'total_number', "proportion"))

df %>%
  group_by(O.ring) %>%
  summarise(mean_Temp = mean(Temp), 
            max_Temp = max(Temp),
            min_Temp = min(Temp),
            median_Temp = median(Temp),
            mean_Pressure = mean(Pressure), 
            max_Pressure = max(Pressure),
            min_Pressure = min(Pressure),
            median_Pressure = median(Pressure),
            n = n()
           )

df %>% group_by(O.ring, Temp) %>% 
  summarise(mean_Pressure = mean(Pressure), 
            max_Pressure = max(Pressure),
            min_Pressure = min(Pressure),
            median_Pressure = median(Pressure),
            count = n())
```
a. There are 23 records with 5 columns.
b. There is no missing data in any columns.
c. We see two scenarios in o-ring failure, either there is a failure or not.
d. The column "Number" denotes total numnber of o-rings present, which is 
constant with value 6.
e. The other columns "Temp" and "Pressure" are explanatory variables.
f. The data set is highly imbalanced. Most of the records are having 0 o-ring
failure type (70%), that is there is no failure followed by 1 incident and 2 
incidents of o-ring failure.

```{r, Histogram}
hist(df$Temp)
hist(df$Pressure)
```
```{r, Corelation}
df.cor = cor(df[c('O.ring', 'Temp', 'Pressure')])
corrplot(df.cor)
```

We see negative correlation between *O.ring* and *Temp* and 
positive correlation between *O.ring* and *Pressure*. 
No correlation was found between *Temp* and *Pressure*. 

```{r, Boxplot, Outlier}
boxplot(df$Temp, main = "Launch temperature")
outlier.Temp <- df %>% filter(Temp < 55)
outlier.Temp
```

We see 2 incidents of -ring failure when the temperature is below 55 deg F.

```{r, Univariate analysis}
oring.plot <- ggplot(df, 
                     aes(x = O.ring)) +
                     geom_histogram(aes(y = ..density..), 
                                    binwidth = 1, 
                                    fill="#0072B2", 
                                    colour="black") +
                     ggtitle("O-ring failure") + 
                     theme(plot.title = element_text(lineheight=1, 
                                                     face="bold"))

Temp.plot <- ggplot(df, 
                    aes(x = Temp)) +
                    geom_histogram(aes(y = ..density..), 
                                   binwidth = 1, 
                                   fill="#0072B2", 
                                   colour="black") +
                    ggtitle("Launch temperature") + 
                    theme(plot.title = element_text(lineheight=1, 
                                                    face="bold"))

Pressure.plot <- ggplot(df, 
                        aes(x = Pressure)) +
                        geom_histogram(aes(y = ..density..), 
                        binwidth = 1, 
                        fill="#0072B2", 
                        colour="black") +
                        ggtitle("Pressure") + 
                        theme(plot.title = element_text(lineheight=1, 
                                                        face="bold"))

grid.arrange(oring.plot, Temp.plot, Pressure.plot, ncol=3)
```
```{r Biovariate analysis between O.ring and temparature}
ggplot(df, aes(factor(O.ring), Temp)) +
  geom_boxplot(aes(fill = factor(O.ring))) + 
  geom_jitter() +
  ggtitle("Launch temperature by O-ring failures") + 
  theme(plot.title = element_text(lineheight=1, face="bold"))
```
We grouped launch temperature by number of O-ring failure incidents and plotted 
each group using boxplot. 
We see that there is no O-ring failure when the flights which were launched 
at higher temperature. We notice that when the temperature drops, it increases
the chances of o-ring failures.

# Trivariate analysis
```{r}
ggplot(df, 
       aes(factor(O.ring), Temp)) +
       geom_boxplot(aes(fill = factor(O.ring))) + 
       facet_wrap(~Pressure) +
       ggtitle("Launch temperature by O-ring failures types at different levels of pressure") + 
       theme(plot.title = element_text(lineheight=1, face="bold"))
```

We ran tri-variate analysis by examining launch temperature by O-ring failures 
at three different pressure levels. 
We see negative correlation between *Temp* and *O-ring* with the pressure level 
200 psi. 
With 200 psi level, one 0-ring failure happens when the average temperature is
slightly over 60 deg F and two o-ring failure happens when the average temperature 
is slightly below 70 deg F.
However, the plot at pressure levels of 50 or 100 psi didn't provide much 
useful information due to limited availability of the data. 

To build a binomial logistic regression model, we used derived feature of
O.ring / Number to give us an observed proportion of O-ring failures.

From the exploratory data analysis, we found negative correlation between *Temp* and 
*O-ring* and positive correlation between *Pressure* and *O-ring*.
We want to include both explanatory features in our first logistic regression model 
to explore more.
For a given launch $i$, we denote the probability for an O-ring to fail as $\pi_i$, 
launch temperature as $t_i$ and leak test pressure as $p_i$. 
With that, the model will be formulated as:-

$$
logit \left( \pi_i \right)  = \beta_0 + \beta_1 t_i + \beta_2 p_i
$$

```{r, model 1}
model_1 <- glm(formula = O.ring / Number ~ Temp + Pressure, 
               data = df,
               family = binomial(link = logit), 
               weights = Number)
summary(model_1)
cbind(Estimate = coef(model_1), confint(model_1))
Anova(model_1, test = 'LR')
```

The results of the likelihood ratio test, we see that temperature appears to be 
statistically significant predictor on the number of o-ring failures. It is also 
showing the inverse relationship between o-ring failure and the temperature. 
The pressure variable is not statistically statistically significant predictor 
on the number of o-ring failures. 
The coefficient of $t_i$ was estimated to be 
`r cbind(Estimate = coef(model_1), confint(model_1))[2,1` with the 95% Wald 
confidence interval of `r cbind(Estimate = coef(model_1), confint(model_1))[2,2]` 
to `r cbind(Estimate = coef(model_1), confint(model_1))[2,3]`, indicating that 
the decrease on the launch temperature would cause the increase on the odds 
for an O-ring to fail. 

In Anova test, involving feature variable of *Temp*, according to 
$H_0 : \beta_1 = 0$ and $H_\alpha : \beta_1 \neq 0$. 
The LRT statistic for *Temp* is `r Anova(model_1, test = 'LR')[1,1]` with a 
p-value of `r Anova(model_1, test = 'LR')[1,3]`. 
Using the Type I Error rate $alpha = 0.05$, we reject the null hypothesis and 
accept that the feature *Temp* is important to be included in the model with
other feature variable *Pressure* is included the model. 
For the test of *Pressure* with $H_0 : \beta_2 = 0$ vs. $H_\alpha : \beta_2 \neq 0$, 
The LRT statistic of *Pressure* is `r Anova(model_1, test = 'LR')[2,1]` with a 
p-value of `r Anova(model_1, test = 'LR')[2,3]`. Using the Type I Error rate 
$alpha = 0.05$, we fail to reject the null hypothesis. 
Therefore, there is a lack of evidence to include the feature variable *Pressure* 
to be included in the model, given that *Temp* is in the model. 

```{r}
test.temp <- -10
exp(test.temp*coef(model_1)['Temp'])
```

With 10 deg F temperature drop increases the odds o-ring failure by approximately 
`r abs(round(floor(exp(test.temp*coef(model_1)['Temp'])) - exp(test.temp*coef(model_1)['Temp']), 2)) * 100`%. 

In our next model, we are just including temperature as our explanatory varaible.
For a given launch $i$, we denote the probability for an O-ring to fail as $\pi_i$, 
launch temperature as $t_i$ . 
With that, the model will be formulated as:-

$$
logit \left( \pi_i \right)  = \beta_0 + \beta_1 t_i
$$

```{r, model 2}
model_2 <- glm(formula = O.ring / Number ~ Temp, 
               data = df,
               family = binomial(link = logit), 
               weights = Number)
summary(model_2)
cbind(Estimate = coef(model_2), confint(model_2))
Anova(model_2, test = 'LR')
```

In our second model, we just used temperature as the feature to predict the
probability of o-ring failure and it still appears to be statistically significant 
feature in predicting the probability of o-ring failure.

Dalal, Fowlkes, and Hoadley (1989) chose to remove *Pressure* from the model 
for the weak effect of it, but we also have limited data for 50 and 100 psi. 
And the interaction between *Temp* and *Pressure* was not taken into account by 
the authors. Given the fact that high pressure can cause "blow holes" in the putty, 
resulting the hot gasses to escape. If so, the effect of pressure can be 
significant predictor when the temperature is very low. Therefore, 
removing *Pressure* from the model could possibly cause serious information loss.

In our third logistic regression model, we use temperature and it's quadratic
term as explanatory features to predict the o-ring failure probability.
For a given launch $i$, we denote the probability for an O-ring to fail as $\pi_i$, 
launch temperature as $t_i$. 
With that, the model will be formulated as:-

$$
logit \left( \pi_i \right)  = \beta_0 + \beta_1 t_i + \beta_2 t_i ^ 2
$$

```{r, model 3}
model_3 <- glm(formula = O.ring / Number ~ Temp + I(Temp ^ 2), 
               data = df,
               family = binomial(link = logit), 
               weights = Number)
summary(model_3)
cbind(Estimate = coef(model_3), confint(model_3))
Anova(model_3, test = 'LR')
```
Then, we wanted to see if the quadratic temperature has any effect on predicting
o-ring failure and we do not see any statistical significance of it. So we are
not going to use this quadratic temperature feature in our model.

In our fourth logistic regression model, For a given launch $i$, we denote the 
probability for an O-ring to fail as $\pi_i$, launch temperature as $t_i$ and 
leak test pressure as $p_i$. We also consider the interaction term between
*Temp* and *Pressure*.
With that, the model will be formulated as:-

$$
logit \left( \pi_i \right)  = \beta_0 + \beta_1 t_i + \beta_2 p_i + \beta_4 t_i:p_i
$$

```{r, model 4}
model_4 <- glm(formula = O.ring / Number ~ Temp + 
                                           Pressure + 
                                           Temp:Pressure, 
               data = df,
               family = binomial(link = logit), 
               weights = Number)
summary(model_4)
cbind(Estimate = coef(model_4), confint(model_4))
Anova(model_4, test = 'LR')
```


```{r}
temp_range <- data.frame(Temp = 31:81)
prob_pred <- predict(model_2, temp_range, type="response", se = TRUE)
temp_range['pi_hat'] <- prob_pred$fit
linear_pred <- predict(model_2, temp_range, type="link", se = TRUE)
lower <- linear_pred$fit - qnorm(1 - 0.05 / 2) * linear_pred$se.fit
upper <- linear_pred$fit + qnorm(1 - 0.05 / 2) * linear_pred$se.fit
temp_range['lower'] <- exp(lower)/(1 + exp(lower))
temp_range['upper'] <- exp(upper)/(1 + exp(upper))

p7 <- ggplot(temp_range, aes(x = Temp, y = pi_hat)) + 
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  labs(title = "Predicted Probability of O-Ring Failure vs Temperature") +
  theme(plot.title = element_text(size = 8), text = element_text(size = 7)) +
  ylab("Pi") +
  xlab("Temperature")

# Expected number of failures vs Temp plot.
temp_range['pred_failures'] <- temp_range$pi_hat * 6
temp_range['failures_lower'] <- temp_range$lower * 6
temp_range['failures_upper'] <- temp_range$upper * 6

p8 <- ggplot(temp_range, aes(x = Temp, y = pred_failures)) + 
  geom_line() + 
  geom_ribbon(aes(ymin = failures_lower, ymax = failures_upper), alpha = 0.2) + 
  labs(title = "Predicted Number of O-Ring Failures vs Temperature") +
  theme(plot.title = element_text(size = 8), text = element_text(size = 7)) +
  ylab("O-Ring Failures") +
  xlab("Temperature")

grid.arrange(p7, p8, ncol=2)
```

The confidence intervals gets wider in lower temperature for lack of training data 
at low temperature, leading the model to be more uncertain at lower temperatures.

When the temperature is 31째 F, the estimated probability of o-ring failure is 
`r round(temp_range[temp_range$Temp == 31,]['pi_hat'], 4)` with the confidence 
interval between `r round(temp_range[temp_range$Temp == 31,]['lower'], 4)` and
`r round(temp_range[temp_range$Temp == 31,]['upper'], 4)`). The estimated number 
of o-ring failures at 31 deg F is `r round(temp_range[temp_range$Temp == 31,]['pred_failures'], 4)` 
with the confidence interval between `r round(temp_range[temp_range$Temp == 31,]['failures_lower'], 4)`
and `r round(temp_range[temp_range$Temp == 31,]['failures_upper'], 4)`. 
In order to draw inference from the model, we need to assume the model is built 
off a representative data set including records with both high and low 
temperatures, high and low pressure levels. We should not be training the model
on a range and use for inference on another range. Then the results will be 
uncertain and we can not trust the model's predictions.

```{r bootstrap-confidence-intervals}
o.ring.failure.estimator <- function(iterations, seed=42, number=6) {
  # Setting seed to make the plot exactly reproduced
  set.seed(seed)
  n <- 23 # Resample size
  estimation <- array(data = NA, dim=c(91, iterations))
  
  # Performing re-sampling repeatedly "b" times
  for(i in 1:iterations) {
    samples <- sample(x = 1:n, size = n, replace = TRUE)
    model   <- glm(formula = O.ring / Number ~ Temp, data = df[samples,],
                   weights = Number, family = 'binomial')
    # Performing prediction for 10-100째 Fahrenheit using each fitted model 
    preds <- predict(object = model, newdata = data.frame(Temp = 10:100),
                     type = "response")
    j <- 0
    for(p in preds) {
      j <- j+1 # Index
      estimation[j, i] <- p
    }
  }
  # Fitting a model with the original data for prediction
  model <- glm(formula = O.ring/Number ~ Temp, data = df,
               weights = Number, family = 'binomial')
  # Estimating Expected number of incidents along with 90% CI
  incidents <- data.frame(Temperature = integer(), Incidents = double(),
                          CI.Upper = double(), CI.Lower = double())
  preds <- predict(object = model, newdata = data.frame(Temp = 10:100),
                   type = "response")
  i <- 0
  for(p in preds) {
    i <- i+1 # Index to locate data in the data.frame
    # Getting the 90% CI and multiply by 6 to get estimation for a single flight
    ci <- quantile(estimation[i,], probs = c(0.05, 0.95), names=FALSE)
    incidents[i, 1] <- i + 9 # Temperature
    incidents[i, 2] <- p*number
    incidents[i, 3] <- ci[1]*number
    incidents[i, 4] <- ci[2]*number
  }

  return(incidents)
}

# Performing predictions for 10째-100째 temperature and plotting the result
predicted.incidents <- o.ring.failure.estimator(2000)
```

```{r}
ggplot(predicted.incidents, aes(x=Temperature)) + 
  geom_line(aes(y = Incidents)) +
  geom_ribbon(aes(ymin = CI.Lower, ymax = CI.Upper), alpha = 0.2) + 
  labs(title = "Expected Number of Incidents by Temperature") +
  theme(plot.title = element_text(size = 10), text = element_text(size = 8)) +
  xlab("Temperature in degree fahrenheit") +
  ylab("Expected number of incidents") -> p9

# Using grid to make the plot smaller to take up less room.
grid.arrange(p9)
```

We are simulating the sampling(with replacement because of small data size) 
process over iterations and performed model fitting using the sample, generated 
in each iteration and used it for prediction and calculation of confidence
interval for different temperature values.
We are running the simulation 2,000 times to generate a stable distribution. 
The confidence interval varies depending on the temperature value. At lower 
temperature, the confidence interval tends to get wider as it moves further away 
from the center of the data, resulting the uncertainty of the model predictions.

```{r, linear model}
model.linear <- lm(O.ring ~ Temp, data = df)
summary(model.linear)

par(mfrow=c(2,2))
plot(model.linear)
```